---
title: "Kuruka"
output: rmarkdown::github_document
editor_options: 
  markdown: 
    wrap: sentence
    author: Riccardo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(svglite))
theme_set(theme_bw())
```

```{}
find /mnt/data2/riccardo/dmel-shortreads/ -name "*.fastq.gz" | parallel -j 5 'n={/.}; bwa bwasw -t 10 ref/c8-derecta_scg.fasta {} | samtools sort -@ 10 -m 3G -o mapped/${n}.sort.bam'

ls -1 *.bam | parallel -j 40 deviaTE_analyse --input {} --single_copy_genes JAEIGS010000001.1:5141171-5166026,JAEIGS010000021.1:2250845-2261726,JAEIGS010000021.1:11118637-11106239 --library ../../ref/kuruka-SCGs.fasta --family "Kuruka"
```

```{}
# Define the folder containing the copynumber files
folder_path <- "/Volumes/Storage/kuruka/deviaTE/raw-data/"

columns <- c("TEfam", "sample_id", "pos", "refbase", "A", "C", "G", "T", "cov", "phys_cov", "hq_cov", "snp", "refsnp", "int_del", "int_del_freq", "trunc_left", "trunc_right", "ins", "delet", "annotation")

# List all *.ori.out files in the folder
file_list <- list.files(path = folder_path, pattern = ".Kuruka", full.names = TRUE)

# Iterate over each file in oriout_folder
all_cn_data <- tibble()

for (cn_file in file_list) {
  cn <- read_table(cn_file, col_names = columns, skip = 3) %>% mutate(sample = cn_file)
  print(cn_file)
  all_cn_data <- bind_rows(all_cn_data, cn)
  }
  
write_tsv(all_cn_data, "/Volumes/Storage/kuruka/deviaTE/kuruka-deviate.tsv")
```

```{r}
metadata <- read_tsv("/Volumes/Storage/kuruka/metadata/dmel-metadata-DEST2.tsv")
```

```{}
read_deviate <- function(path, meta){
  te <- read_tsv(path) %>% select(TEfam, sample_id, pos, A, C, G, T, cov, hq_cov, snp) %>% mutate(sample_id = gsub(".fastq.sort.bam", "", sample_id)) %>% inner_join(meta, by = "sample_id") %>% select(-study, -study_id, -publication)
}

kuruka <- read_deviate("/Volumes/Storage/kuruka/deviaTE/kuruka-deviate.tsv", metadata)
```

```{}
# Define the folder containing the copynumber files
folder_path <- "/Volumes/Storage/kuruka/deviaTE/raw-data/"

columns <- c("TEfam", "sample_id", "pos", "refbase", "A", "C", "G", "T", "cov", "phys_cov", "hq_cov", "snp", "refsnp", "int_del", "int_del_freq", "trunc_left", "trunc_right", "ins", "delet", "annotation")

# List all *.ori.out files in the folder
file_list <- list.files(path = folder_path, pattern = ".Kuruka", full.names = TRUE)

# Iterate over each file in oriout_folder
all_cn_data <- tibble()

for (cn_file in file_list) {
  cn <- read_table(cn_file, col_names = columns, skip = 3) %>% mutate(sample = cn_file) %>% filter(pos < 7200 | pos > 7275) %>% group_by(sample) %>% summarise(copynumber = mean(hq_cov), more_than1 = sum(hq_cov >= 1)) %>% mutate(presence = ifelse(more_than1 > 7005, "present", "absent"))
  all_cn_data <- bind_rows(all_cn_data, cn)
  }

print(all_cn_data)
write_tsv(all_cn_data, "/Volumes/Storage/kuruka/deviaTE/kuruka-copynumber-nolowcomplexity.tsv")
```

7200-7275 <- low complexity region with high copynumber

```{r}
copynumber_kuruka <- read_tsv("/Volumes/Storage/kuruka/deviaTE/kuruka-copynumber-nolowcomplexity.tsv") %>% mutate(sample = gsub("/Volumes/Storage/kuruka/deviaTE/raw-data//", "", sample), sample = gsub(".fastq.sort.bam.Kuruka", "", sample))

period_order <- c("<2010", "2010-2017", "2017-2019", "2020-2021")

(metadata_copynumber <- inner_join(metadata, copynumber_kuruka, by="sample") %>% mutate(period = case_when(year<2010 ~ "<2010", year>=2010 & year<2017 ~ "2010-2017", year>=2017 & year < 2020 ~ "2017-2019", TRUE ~ "2020-2021"), period = factor(period, levels = period_order)) %>% arrange(desc(copynumber)))

world_map <- map_data("world") %>% filter(region != "Antarctica")

(kuruka_timeline <- ggplot(metadata_copynumber, aes(x = year, y = copynumber, color = presence)) +
        geom_point(alpha = 0.3, size = 3) +
        labs(y = "copynumber", x = "", color = "Kuruka") +
        #scale_color_gradient(low = "darkgreen", high = "red") +
        scale_color_manual(values = c("darkgreen", "red")) +
        theme(legend.position = "top"))

(c8_map <- ggplot() +
        geom_map(data = world_map, map = world_map, 
                 aes(long, lat, map_id = region), 
                 fill = "cornsilk2", color = "darkgrey", linewidth = 0.1) +
        geom_point(data = metadata_copynumber, 
                   aes(x = lon, y = lat, color = presence), 
                  size = 2, position = position_jitter(width = 2, height = 2), alpha = 0.3) +
        #scale_color_gradientn(colours = c("darkgreen", "red", "darkred")) +
    scale_color_manual(values = c("darkgreen", "red")) +
       theme(plot.title = element_text(hjust = 0.5), 
              axis.text = element_blank(), 
              axis.title = element_blank(), 
              axis.ticks = element_blank(), 
              legend.position = "bottom", 
              panel.grid = element_blank(), 
             panel.background = element_rect(fill = "lightblue")) +
        facet_wrap(~period) +
        labs(color = "Kuruka") +
       coord_cartesian(expand = FALSE))

ggsave("/Volumes/Storage/kuruka/figures/map.png", c8_map)
ggsave("/Volumes/Storage/kuruka/figures/timeline.png", kuruka_timeline, height = 2, dpi = 600)
write_tsv(metadata_copynumber, "/Volumes/Storage/kuruka/deviaTE/kuruka-copynumber-nolowcomplexity-metadata.tsv")
```

```{r}
(c8_timeline_zoom <- ggplot(metadata_copynumber, aes(x = year, y = copynumber, color = presence)) +
        geom_point(alpha = 0.5, size = 3) +
        labs(y = "copynumber", x = "", color = "Kuruka") +
        scale_color_manual(values = c("darkgreen", "red")) +
        theme(legend.position = "top") +
        xlim(2000, 2021))

c8_africa <- metadata_copynumber %>% filter(lat > -45 & lat < 40, lon > -20 & lon < 55)
c8_europe <- metadata_copynumber %>% filter(lat > 36 & lat < 70, lon > -10 & lon < 40)
c8_asia <- metadata_copynumber %>% filter(lat > 10 & lat < 75, lon > 80 & lon < 140) %>% mutate(period = case_when(year==2017 ~ "2017", year==2020 ~ "2020", TRUE~"<2017"))

naeu_map <- world_map %>% filter(lat > 25 & lat < 70, long > -130 & long < 45)
europe_map <- world_map %>% filter(lat > 36 & lat < 70, long > -10 & long < 40)
africa_map <- world_map %>% filter(lat > -45 & lat < 40, long > -20 & long < 55)
asia_map <- world_map %>% filter(lat > 10 & lat < 75, long > 80 & long < 140)

  
(c8_europe_map <- ggplot() +
        geom_map(data = europe_map, map = europe_map, 
                 aes(long, lat, map_id = region), 
                 fill = "cornsilk2", color = "darkgrey", linewidth = 0.1) +
        geom_point(data = c8_europe, 
                   aes(x = lon, y = lat, color = presence), 
                   size = 3, position = position_jitter(width = 2, height = 2), alpha = 0.4) +
        scale_color_manual(values = c("darkgreen", "red")) +
        #scale_color_gradientn(colours = c("darkgreen", "red", "darkred")) +
        theme(plot.title = element_text(hjust = 0.5), 
              axis.text = element_blank(), 
              axis.title = element_blank(), 
              axis.ticks = element_blank(), 
              legend.position = "bottom", 
              panel.grid = element_blank(), 
              panel.background = element_rect(fill = "lightblue")) +
        facet_wrap(~period) +
        labs(color = "Kuruka") +
        coord_cartesian(expand = FALSE))

(c8_asia_map <- ggplot() +
        geom_map(data = asia_map, map = asia_map, 
                 aes(long, lat, map_id = region), 
                 fill = "cornsilk2", color = "darkgrey", linewidth = 0.1) +
        geom_point(data = c8_asia, 
                   aes(x = lon, y = lat, color = presence), 
                   size = 3, position = position_jitter(width = 2, height = 2), alpha = 0.4) +
        #scale_color_gradientn(colours = c("darkgreen", "red", "darkred")) +
        scale_color_manual(values = c("darkgreen", "red")) +
        theme(plot.title = element_text(hjust = 0.5), 
              axis.text = element_blank(), 
              axis.title = element_blank(), 
              axis.ticks = element_blank(), 
              legend.position = "bottom", 
              panel.grid = element_blank(), 
              panel.background = element_rect(fill = "lightblue")) +
        facet_wrap(~period) +
        labs(color = "Kuruka") +
        coord_cartesian(expand = FALSE))

ggsave("/Volumes/Storage/kuruka/figures/map-europe.png", c8_europe_map)
ggsave("/Volumes/Storage/kuruka/figures/map-asia.png", c8_asia_map)
ggsave("/Volumes/Storage/kuruka/figures/timeline-recent.png", c8_timeline_zoom, height = 2, width = 5, dpi = 600)
```

```{}
deviaTE_analyse --input SRR23103981.fastq.sort.bam --single_copy_genes JAEIGS010000021.1:3993639-3984763,JAEIGS010000049.1:3568949-3579275,JAEIGS010000103.1:16029183-16016880 --library ../ref/c8-derecta_scg.fasta --family c8_derecta
```
